The force-free magnetosphere of a Kerr black hole: the role of boundary conditions

### 2018-06-26 paper/ms.tex
state save

\begin{document}
\title{The force-free magnetosphere of a Kerr black hole: the role of boundary conditions}
\author{Zhen Pan}
\email{zhpan@ucdavis.edu}
\affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario N2L 2Y5, Canada}

@@

@@

\section{Uniform field solution}

@@

The boundary conditions
\begin{subequations}
\begin{align}
    A_\phi(\mu = 1) &= 0,  \label{eq:bc1}\\
    \Am(\mu = 0, r > 2) &= 0, \label{eq:bc2}\\
    B^2-E^2 (\mu = 0, r_+ \leq r \leq 2) &=0.\label{eq:bc3}
\end{align}
\end{subequations}
where $r_+$ is the radius of the event horizon.

In our computation, we use following numerical prescription
\be
\label{eq:nbc}
\int_{A_\phi^{\rm HE}}^{A_\phi^{\rm EE}} \frac{B^2-E^2}{B^2+E^2} dA_\phi \Bigg/ \left(A_\phi^{\rm EE}- A_\phi^{\rm HE}\right)  < 10^{-3}.
\ee
as a proxy of the marginally force-free equatorial boundary condition (\ref{eq:bc3}),
where ``HE" and ``EE" are short for Horizon-Equator and Ergosphere-Equator, respectively;
and we choose $B^2+E^2$ to be the energy density measured by the zero-angular-momentum-observers.
Specifically,
\be
\begin{aligned}
B^2-E^2 &= \frac{1}{\Sigma \sst} \left[ -\mathcal{K} \left(\Ar^2 +\frac{\sst}{\Delta}\Am^2 \right)+\frac{\Sigma}{\Delta}I^2\right]\\
B^2+E^2 &= \frac{1}{\Sigma \sst} \left[ \left(\mathcal{K}+\frac{\Delta\Sigma}{\beta} \right) \left(\Ar^2 +\frac{\sst}{\Delta}\Am^2 \right)+\frac{\Sigma}{\Delta}I^2\right]\\
\end{aligned}
\ee

\subsection{Numerical method}
We aim to find a pair of $\Omega(A_\phi)$ and $I(A_\phi)$ connected by the radiation condition
\be I = 2\Omega A_\phi, \label{eq:rad}\ee
and enabling field lines smoothly crossing the LS,
and suitable normal derivative $\Am(\mu =0, r_+\leq r \leq 2)$ on the equator
guaranteeing the boundary condition (\ref{eq:bc3}).

In our computation, we define a new radial coordinate $R=r/(1+r)$, confine our
computation domain $R\times \mu$ in the region $[R(r_+), 1]\times [0,1]$,
and implement a uniform $512\times 64$ grid. The detailed algorithm is clarified
in the following steps.

@@

4. The remaining task is to adjust  $\Omega(0< \Ap < \AHE)$ enabling  the radiation condition (\ref{eq:rad})
 for $A_\phi \in (0, A_\phi^{\rm HE})$ and  to adjust $\Omega(\AHE \leq \Ap \leq \AEE)$
 enabling the boundary condition (\ref{eq:bc3}) for $\Ap \in [\AHE, \AEE]$.
 The first part is straightforward, i.e.,
 \be
2\Ap\Omega_{\rm new}  = I|_{0 < \Ap < \AHE},
 \ee
 and the second part can be realized by iterative correction
 \be
2\Ap(\Omega_{\rm new} - \Omega_{\rm old}) =  - \zeta_2\times\Delta (B^2-E^2)|_{\mu=0, r_+\leq r \leq 2},
 \ee
 where $\zeta_2$ is again an empirical step size, and we have multiplied factor $\Delta$
 in the correction term to avoid numerical difficulty in the vicinity of the event horizon.
 To eliminate unphysical discontinuity in the angular
 velocity at $\AHE$, we fit $\Omega_{\rm new}(\Ap)$ on the whole range
 $(0, \AEE)$ via a fifth-order polynomial.

@@

In Figure \ref{fig:field_lines}, we plot the magnetic field lines enclosing a BH with spin $a =0.99$ as an example,
and in the upper panel of Figure \ref{fig:omega}, we show the angular velocity functions $\Omega(A_\phi)$ for different BH spins.
Combining the two figures, we see three generic features: (1) the LS runs from $r=r_+$ to $r=2M$ as $\theta$ varies from $0$ to $\pi/2$,
(2) there is no current sheet within the magnetosphere except the equatorial current sheet extending from $r_+$ to $2M$,
which gives rise to a cusp ($\Am \neq 0$) to the equatorial magnetic field lines,
(3) magnetic field lines entering the ergosphere end up either on the horizon or on the equatorial current sheet,
both of which carry electric current and therefore Poynting flux.

In the lower panel of Figure  \ref{fig:omega} and in Figure \ref{fig:Edot}, we compare our numerical results with
recent high-resolution simulations performed by East and Yang \cite{East2018} (EY Sims). In Figure \ref{fig:Edot}, the solid
lines are curves fitted by polynomials

We find the angular velocity functions
in two studies show a good agreement, while the value of our magnetic flux $\AEE$ is lower by $\sim 5\%$, which gives
to $\sim 10\%$ lower energy extraction rates.

@@

\begin{figure}
\includegraphics[scale=0.6]{f3}
\caption{\label{fig:Edot} Comparison of the energy extraction rate $\dot E(\WH)$ derived from our
numerical solutions and from the simulations \cite{East2018},
where the former is lower than the latter one by $~10\%$.}
\end{figure}

\section{Discussions}

\subsection{Solution uniqueness}

\subsection{Topology of near-horizon field lines}

\section{Summary}

>>>>

\begin{document}
\title{The force-free magnetosphere of a Kerr black hole: the role of boundary conditions}
\author{Zhen Pan}
\email{zpan@perimeterinstitute.ca}
\affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario N2L 2Y5, Canada}

@@

The paper is organized as follows.  In Section \ref{sec:basic}, we outline the basic governing equations.
In Section \ref{sec:uni_sol}, we clarify the radiation condition, boundary conditions and numerical algorithm
for the uniform field solution. In Section \ref{sec:discussion}, we discuss the solution uniqueness, near-horizon
field line configuration and the application to general field line configurations. Summary is given in Section \ref{sec:summary}.

@@

\section{Uniform field solution}
\label{sec:uni_sol}

\subsection{Solution uniqueness and radiation condition}
For common field configurations, including split monopole and parabola, there exist two LSs where
the LS function vanishes and the GS equation degrades from second order to first order. As proposed
by Contopoulos et.al. \cite{Contopoulos2013}, one can adjust the two free functions $\Omega(\Ap)$
and $I(\Ap)$ enabling field lines smoothly cross the two LSs, then the solution
$\{\Omega(\Ap), I(\Ap), \Ap(r,\theta)\}$  is uniquely found. But for the vertical field lines, their
exist only one LS, which is insufficient to determine two free functions. Therefore people have been
wondering the possibility of many solutions though the many-solutions scenario is not supported by force-free
dynamics simulations. To tackle this problem, we proposed that the two functions are not independent;
instead, they are related by the radiation condition \cite{Pan2016a, Pan2017}
\be I = 2\Omega A_\phi, \label{eq:rad}\ee
which has been confirmed by the recent high-accuracy simulations \cite{East2018}.

@@

The boundary conditions at infinity (inner infinity $r=r_+$ and outer infinity $r\rightarrow\infty$)
and on the polar axis can be simply set as
\be
\begin{aligned}
\Ar|_{r=r_+, \infty} &= 0, \\
A_\phi|_{\mu = 1} &= 0,\\
\end{aligned}
\ee
where $r_+$ is the radius of the event horizon,
while the equatorial boundary conditions are more unclear until recent simulations \cite{East2018}
come out which show that there exists an equator current sheet within the ergosphere where the
magnetic dominance marginally loses, i.e., $B^2-E^2\approx 0$. Motivated by these simulations,
we choose the following equatorial boundary condition in our numerical solutions,
\begin{subequations}
\begin{align}
    \Am(\mu = 0, r > 2) &= 0, \label{eq:bc2}\\
    B^2-E^2 (\mu = 0, r_+ \leq r \leq 2) &=0.\label{eq:bc3}
\end{align}
\end{subequations}
In fact, Equation (\ref{eq:bc3})
is neither a Dirichlet nor a Neumann boundary condition, since
\be
\begin{aligned}
&B^2-E^2 \\
= & \frac{1}{\Sigma \sst} \left[ -\mathcal{K} \left(\Ar^2 +\frac{\sst}{\Delta}\Am^2 \right)+\frac{\Sigma}{\Delta}I^2\right],
\end{aligned}
\ee
which involves both derivatives $\Am$ and $\Ar$ on the boundary.
Therefore, it is numerically non-trivial to impose this boundary condition in computation.

In addition, there is a coordinate singularity $1/\Delta$ in the expression of $B^2-E^2$.
To avoid possible numerical difficulty, we use following prescription
\be
\label{eq:nbc}
\int_{A_\phi^{\rm HE}}^{A_\phi^{\rm EE}} \left(\frac{B^2-E^2}{B^2+E^2}\right)^2 dA_\phi \Bigg/ \left(A_\phi^{\rm EE}- A_\phi^{\rm HE}\right)  < 10^{-3}.
\ee
in our computation, as a proxy of the marginally force-free equatorial boundary condition (\ref{eq:bc3}),
where $\AHE$ and $\AEE$ are the magnetic flux entering the horizon and the ergosphere, respectively;
``HE" and ``EE" are short for Horizon-Equator and Ergosphere-Equator, respectively.
For definiteness, we choose $B^2+E^2$ to be the energy density measured by the zero-angular-momentum-observers.
Explicitly, we have
\be
\begin{aligned}
 & B^2+E^2   \\
=& \frac{1}{\Sigma \sst} \left[ \left(\mathcal{K}+\frac{\Delta\Sigma}{\beta} \right)
 \left(\Ar^2 +\frac{\sst}{\Delta}\Am^2 \right)
 +\frac{\Sigma}{\Delta}I^2\right]
\end{aligned}
\ee

\subsection{Generic features of the BH magnetosphere structure}
Before delving into the details of numerically solving the GS equations,
we point out that from the radiation condition (\ref{eq:rad}) and the marginally
force-free boundary condition (\ref{eq:bc3})
themselves contains rich information about the BH magnetosphere structure.

Let's first find out where the LS intersect with the equator, $r_{\rm LS}|_{\mu=0}$.
From the expression of the LS function, it is straightforward to see $ r_{\rm LS}|_{\mu=0} \leq 2$ for $\Omega \geq 0$.
At this point,  the LS function $\mathcal K$ vanishes,
and $I$ must also vanish for satisfying the boundary condition (\ref{eq:bc3}),
which in turns indicates a vanishing angular velocity $\Omega$ from
the radiation condition (\ref{eq:rad}).  Due to the static limit, $\Omega$
must be non-zero inside the ergosphere.  Therefore the only possibility is  $r_{\rm LS}|_{\mu=0} = 2$, i.e.
to satisfy the boundary condition (\ref{eq:bc3}), the LS must interact the equator at $r=2$, which also
justifies our choice of equatorial boundary conditions (\ref{eq:bc2},\ref{eq:bc3}) .

From above analysis, we expect several generic features in the magnetospheres that are independent
of the GS equation:
(1) the LS runs from $r=r_+$ to $r=2$ as $\theta$ varies from $0$ to $\pi/2$,
(2) there is no current sheet within the magnetosphere except the equatorial current
sheet extending from $r_+$ to $2$, which gives rise to a cusp ($\Am \neq 0$) to the equatorial magnetic field lines,
(3) magnetic field lines entering the ergosphere end up either on the horizon or on the equatorial current sheet,
both of which carry electric current and therefore Poynting flux. With the guidance of the qualitative
features above, we are to numerically solve the GS equation and quantify these features.

\subsection{Numerical method}
We aim to find a pair of $\Omega(A_\phi)$ and $I(A_\phi)$ satisfying the radiation condition (\ref{eq:rad})
and enabling field lines smoothly crossing the LS,
and suitable normal derivative $\Am(\mu =0, r_+\leq r \leq 2)$ on the equator
guaranteeing the boundary condition (\ref{eq:bc3}), which itself cannot be directly
used in computation.

In our computation, we define a new radial coordinate $R=r/(1+r)$, confine our
computation domain $R\times \mu$ in the region $[R(r_+), 1]\times [0,1]$,
and implement a uniform $512\times 64$ grid. The numerical algorithm for
searching desired $\{\Omega(A_\phi), I(A_\phi), \Am(\mu =0, r_+\leq r \leq 2)\}$ is detailed
in the following steps.

@@

4. The remaining task is to adjust  $\Omega(0< \Ap < \AHE)$ enabling  the radiation condition (\ref{eq:rad})
 for $A_\phi \in (0, A_\phi^{\rm HE})$ and  to adjust $\Omega(\AHE \leq \Ap \leq \AEE)$
 enabling the boundary condition (\ref{eq:bc3}) for $\Ap \in [\AHE, \AEE]$.
 The first part is straightforward, i.e.,
 \be
2\Ap\Omega_{\rm new}  = I|_{0 < \Ap < \AHE},
 \ee
 and the second part can be realized by iterative corrections
 \be
2\Ap(\Omega_{\rm new} - \Omega_{\rm old}) =  - \zeta_2\times\Delta (B^2-E^2)|_{\mu=0, r_+\leq r \leq 2},
 \ee
 where $\zeta_2$ is again an empirical step size, and we have multiplied factor $\Delta$
 in the correction term to avoid numerical difficulty in the vicinity of the event horizon.
 To eliminate unphysical discontinuity in the angular
 velocity at $\AHE$, we fit $\Omega_{\rm new}(\Ap)$ on the whole range
 $(0, \AEE)$ via a fifth-order polynomial.

@@

In Figure \ref{fig:field_lines}, we plot the magnetic field lines enclosing a BH with spin $a =0.99$ as an example,
and in the upper panel of Figure \ref{fig:omega}, we show the angular velocity functions $\Omega(A_\phi)$ for different BH spins.

@@

\begin{figure}
\includegraphics[scale=0.6]{f3}
\caption{\label{fig:Edot} Comparison of the energy extraction rate $\dot E(\WH)$ obtained from our
numerical solutions and from the high-resolution force-free simulations \cite{East2018},
where the former is lower than the latter by $\sim 10\%$ for large spins.}
\end{figure}

With the angular velocity $\Omega(\Ap)$ available, the energy extraction rate from the BH is known as
\be
\label{eq:Edot}
\dot E = 4\pi \int_0^{\AEE} \Omega \times I \ d \Ap.
\ee
In the limit of slow rotation, analytic solution is known as \cite{Pan2014, East2018},
\be
    \Omega = \WH\frac{\sqrt{1-\psi}} {1+\sqrt{1-\psi}},
\ee
where $\psi = \Ap/(2B_0M^2)$, and the energy extraction rate is
\be
\label{eq:lowEdot}
\dot E = 128\pi \left(\frac{17}{24}-\ln 2\right) B_0^2 M^4 \WH^2.
\ee

In the lower panel of Figure  \ref{fig:omega} and in Figure \ref{fig:Edot}, we compare our numerical results with
recent high-resolution simulations done by East and Yang \cite{East2018}. In Figure \ref{fig:Edot},
the points are taken from either the simulations or our numerical solutions, while the solid lines are
corresponding polynomial fitting curves which require to approach Eq.(\ref{eq:lowEdot}) for small spins and to be flat for extremal spins.
We find the angular velocity functions obtained from two different approaches yield a good agreement (Figure \ref{fig:omega}),
while the energy extraction rate $\dot E$ is $\sim 10\%$ lower than the corresponding simulation results, which is traced to
$\sim 5\%$ smaller magnetic flux $\AEE$ in our numerical solutions.

\section{Discussions}
\label{sec:discussion}

\subsection{General application}
In real astrophysics environment, we expect the field configurations are more close to parabolas instead of
being strictly vertical. Due to the same topology, it is reasonable to expect the marginally force-free boundary
condition (\ref{eq:bc3}) is also a good work approximation. The features we observed the uniform field solution
should also apply here. The numerical method for solving the GS equation and self-consistently imposing the boundary
condition (\ref{eq:bc3}).

\subsection{Near-horizon field lines}
In one of our previous papers \cite{Pan2016a}, we made a claim that ``in the steady axisymmetric force-free magnetosphere
around a Kerr BH, all magnetic field lines that cross the infinite-redshift surface must intersect the event
horizon" based on two
However, simulations shows that there exists an equatorial current sheet within the ergosphere, where
the force-free condition breaks down.

\section{Summary}
\label{sec:summary}

In this paper, we revisit the problem of the uniform field solution to the BZ mechanism.
